---
title: "Secreted and Transmembrane"
author: "Maggie Shaw, Temple University"
date: "5/3/2018"
output:
  html_document: default
  word_document: default
---
### Objective: 
Analyze the database results for each protein (indicated by the corresponding protein coding gene hgnc_symbol) to determine a final comprehensive list of "transmembrane" and "secreted" proteins. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Libraries
library(readr)
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(DataCombine)
library(UpSetR)
require(knitr)
require(purrr)
require(broom)
require(modelr)
require(pROC)
library(biomaRt)


# Read in tsv files with transmembrane/secreted lists
trans.list.scores <- read.csv("trans.list.scores2.tsv", header = TRUE, sep = "\t")
secr.list.scores <- read.csv("secr.list.scores2.tsv", header = TRUE, sep = "\t")

```
<br>


##Summary of each source 
Paired bar plot depicts the number of results (secreted/transmembrane) provided by each source. 
```{r}
# Get counts from data frames

numb.source <- function(mylists){
  # remove last column from dataframe
  mylists$Standard_Protein <- NULL
  as.data.frame(colSums(mylists !=0)) -> my.numbs
  colnames(my.numbs)[1]<- "sums"
  rownames(my.numbs) -> my.numbs$source
  my.numbs %>% spread(source, sums, convert = FALSE) -> my.numbs
  
  my.numbs[ , -which(names(my.numbs) %in% c("hgnc_symbol"))] -> my.numbs

  return(my.numbs)
}

numb.source(trans.list.scores) -> trans.list.numbers
trans.list.numbers[4] <- NULL
colnames(trans.list.numbers) <- c("Compartments", "CSPA", "Gene Ontology", "Gold Standard", "LOCATE", "LocDB","Negative Control", "OrganelleDB" , "Uniprot")
trans.list.numbers <- trans.list.numbers[c("Compartments", "CSPA", "Gene Ontology", "LOCATE", "LocDB", "OrganelleDB" , "Uniprot", "Gold Standard", "Negative Control" )]


numb.source(secr.list.scores) -> secr.list.numbers
secr.list.numbers[3] <- NULL
colnames(secr.list.numbers) <- c("Compartments", "Gene Ontology", "Gold Standard", "LOCATE", "Negative Control", "Uniprot")
secr.list.numbers %>% mutate(CSPA = 0) %>% mutate(LocDB = 0) %>% mutate(OrganelleDB = 0) -> secr.list.numbers
secr.list.numbers <- secr.list.numbers[c("Compartments", "CSPA", "Gene Ontology", "LOCATE", "LocDB", "OrganelleDB" , "Uniprot", "Gold Standard", "Negative Control")]

counts.lists <- function(mycountslist, newcol){
  mycountslist %>% mutate(type = newcol) -> mycountslist
  mycountslist %>% gather(source, counts, `Compartments`:`Negative Control`)-> mycountslist
  
  return(mycountslist)
}

counts.lists(trans.list.numbers, "Transmembrane Counts") -> trans.list.numbers
counts.lists(secr.list.numbers, "Secreted Counts") -> secr.list.numbers
rbind(trans.list.numbers, secr.list.numbers) -> all.counts





pro_barplot <- ggplot(all.counts, aes(source, counts, fill = type)) + geom_bar(stat = "identity", position = "dodge") + xlab("Source") + ylab("Protein Counts") + geom_text(aes(label=counts), vjust=1.6, color="white", position = position_dodge(0.9), size = 2.5) + scale_fill_brewer() + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
                                                                                                                        
print(pro_barplot)

```

<br>


## Upset Plots as a Comparison Figure
Upset plots are used to compare all the overlapping proteins between sources. 

###"Transmembrane"
```{r}
# Rename columns for clarity on the plot

colnames(trans.list.scores) <- c("hgnc_Name", "OrganelleDB", "Compartments",  "CSPA", "LOCATE","LocDB","Gene Ontology" , "Uniprot", "Gold Standard", "Gold Standard.r", "Negative Control")

colnames(secr.list.scores) <- c("hgnc_Name", "Compartments", "LOCATE", "Gene Ontology", "Uniprot", "Gold Standard", "Gold Standard.r", "Negative Control")


# Upset plots with something similar to the bar graph shown above
transup <- upset(trans.list.scores, sets = c("OrganelleDB", "Compartments", "CSPA", "LOCATE", "LocDB","Gene Ontology" , "Uniprot", "Gold Standard", "Negative Control"), keep.order = TRUE, nintersects = 25, cutoff = 50, sets.bar.color = "#79CDCD", mainbar.y.label = "Number of Genes", sets.x.label = "Total Number of Genes", text.scale = c(1.5, 1, .75, .75, 1, 1), group.by = "degree", order.by="freq")



```
<br>


###"Secreted"
```{r}
secrup <- upset(secr.list.scores, sets = c("Compartments", "LOCATE", "Gene Ontology", "Uniprot", "Gold Standard", "Negative Control"), keep.order = TRUE, nintersects = 25, cutoff = 50, sets.bar.color = "#EE799F", mainbar.y.label = "Number of Genes", sets.x.label = "Total Number of Genes", text.scale = c(1.5, 1, .75, .75, 1, 1), group.by = "degree", order.by="freq")


```
<br>




##Sum Scores to Produce Final Lists<br>

Sum the scores from each source (including "gold standard", excluding "negative control"), and only include those which occur in at least half of the sources, to produce the final lists. Remove the other columns except the number of sources the protein occured in and the hgnc_symbol. 

```{r}
# Function to combine columns to create new lists and compare against Gold Standard
# Creat Sum Column
sum.scores <- function(my.df, columns.to.sum){
  my.df %>% mutate(Sum.Scores = rowSums(.[columns.to.sum])) -> my.df

  return(my.df)
}


sum.scores(trans.list.scores, 2:9) -> trans.list.scores
sum.scores(secr.list.scores, 2:6) -> secr.list.scores

# Want to produce a list where protein found in at least half of sources (including gold standard)

trans.list.scores %>% filter(Sum.Scores >= 4) -> trans.list.final
trans.list.final[,c("hgnc_Name", "Sum.Scores")] -> trans.list.final
secr.list.scores %>% filter(Sum.Scores >= 3) -> secr.list.final
secr.list.final[,c("hgnc_Name", "Sum.Scores")] -> secr.list.final


colnames(trans.list.final) <- c("HGNC Symbol", "Number of Databases Confirmed")
colnames(secr.list.final) <- c("HGNC Symbol", "Number of Databases Confirmed")
write.table(trans.list.final, "transmembrane_proteins.tsv", quote = FALSE, sep = '\t')
write.table(secr.list.final, "secreted_proteins.tsv", quote = FALSE, sep = '\t')

print(head(trans.list.final))
print(head(secr.list.final))

```




